---
output:
  word_document: default
  html_document: default
---
# BAN 502 Predictive Analytics

## Module 2, 1st R Assignment

### Matthew Field


```{r load packages and data}
library(tidyverse)
library(tidymodels)
library(GGally)
library(lmtest)
air = airquality
```

### Reviewing Questions

Question 1 How many rows are in the “air” dataframe?
This is a 153x6 dataframe, there are 153 rows.

Question 2 How many columns are in the “air” dataframe?
This is a 153x6 dataframe, there are 6 columns.

Question 3 True/False: There is missing data in “Ozone” variable in the dataframe.
```{r Question 3}
summary(air)
# check for NAs or NAs in the values such as characters.
```
There are 44 NA's in 2 variables.

Question 4 Which variable is most likely to be the response (Y) variable?
```{r Question 4, 5, 6}
head(air)
# wind temp month and day are all not likely response variables, ozone seems most likely
model_1 <- lm(air)
summary(model_1)

air2 = air %>%
  drop_na()
```
Ozone

Question 5 How many rows remain in this new (air2) data frame?
with the rows removed containing NA, there are 111 remaining rows.

Question 6 How many columns remain in this new (air2) data frame?
Same amount of columns, 6

Question 7 Which variable is most strongly correlated with the “Ozone” variable?
```{r Question 7 and 8}
ggpairs(air2) # plots lots of helpful data included correlation to other variables
ggcorr(air2, label = TRUE) #shows quick correlation values
```

Temperature is most strongly correlated.

Question 8 Which variable is least strongly correlated with the “Ozone” variable?
Day is least strongly correlated

Question 9 Plot “Temp” (x axis) versus “Ozone” (y axis) using the “ggplot” function. Choose an appropriate chart type. Which statement best describes the relationship between “Temp” and “Ozone”?

```{r Question 9}
ggplot(air2, aes(x=Temp, y = Ozone)) + 
  geom_point() +
  geom_smooth(method = 'lm')
```

As temperature increases, ozone increases

Question 10 What is the slope of this regression model (to four decimal places)?
```{r Question 10}
ozone_simple = recipe(Ozone ~ Temp, air2)

#Next we specify the type of model that we are building.  
lm_model = #give the model type a name 
  linear_reg() %>% #specify that we are doing linear regression
  set_engine("lm") #specify the specify type of linear tool we want to use 

 
#Next we combine the recipe and the model with a workflow.  

lm_wflow = 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(ozone_simple)

 
#Next we fit (execute) the workflow on our dataset.  

lm_fit = fit(lm_wflow, air2)
summary(lm_fit$fit$fit$fit)
```

The slope for this model is 2.4391

Question 11 what is the R-squared value of this model (not Adjusted R-squared) (to three decimal places)?
The R^2 is 0.488

Question 12 Is the “Temp” variables significant in the model?
Yes because p-value is less than 0.05

Question 13 Use the code below to generate 95% confidence intervals for the coefficients. Note that you may need to change “lm_fit” to the name of your model fit if you used a different name
True/False: A 95% confidence interval for the slope coefficient does not contain zero.

```{r Question 14}
confint(lm_fit$fit$fit$fit)
```



Question 14: Using your linear regression model with “Temp” to predict “Ozone”, what is the predicted
“Ozone” value when “Temp” is equal to 80 (to two decimal places)?
47.48

Question 15 Perform appropriate model diagnostics to verify whether or not the model appears to meet the four linear regression model assumptions.
True/False: There is no evidence of non-independent (autocorrelated) residuals.

The predictor and response variable have a linear relationship, yes, shows linear, and fits relatively well
Model errors (residuals) are independent (recall that a residual is the difference between a predicted value and the actual value)
```{r dw test}
dwtest(lm_fit$fit$fit$fit)
```
passes dw test
Model residuals exhibit constant variance, false
```{r}
air2 = air2 %>% mutate(resid_air2 = lm_fit$fit$fit$fit$residuals) #add the model residuals to our data frame
ggplot(air2,aes(x=sqrt(Temp),y=resid_air2)) + geom_point() + theme_bw()

ggplot(air2,aes(x=resid_air2)) + geom_histogram() + theme_bw()
```

Model residuals are Normally-distributed
They are with some outliers


