# BAN 502 Predictive Analytics

## Module 2 Assigment 2 Multiple Linear Regression (MLR)

### Matthew Field 05/26/25

```{r load information}
library(tidyverse) #tidyverse set of packages and functions
library(tidymodels)
library(glmnet) #for Lasso, ridge, and elastic net models 
library(GGally) #create ggcorr and ggpairs plots
library(ggcorrplot) #create an alternative to ggcorr plots
library(MASS) #access to forward and backward selection algorithms
library(leaps) #best subset selection
library(lmtest) #for the dw test
library(splines) #for nonlinear fitting

bike <- read_csv("bike_cleaned-3.csv")
bike = bike %>%
  mutate(dteday = mdy(dteday)) %>%
  mutate_if(is.character, as_factor) %>%
  mutate(hr = as_factor(hr))
```

Question 1, checking correlation with count

```{r Question 1}
ggcorr(bike, label = TRUE, label_round = 3)
```

Question 1, comparing windspeed, hum, atemp, and temp, temp is highest.

Question 2 Boxplot

```{r Question 2}
ggplot(bike,aes(x=season,y=count)) + 
  geom_boxplot() + 
  geom_jitter(alpha = 0.15) + 
  theme_bw()
```
summer has the most bike rides

```{r Question 3}
bike_recipe = recipe(count ~ hr, bike)
lm_model = 
  linear_reg() %>%
  set_engine("lm")

lm_wflow = 
  workflow() %>%
  add_model(lm_model) %>%
  add_recipe(bike_recipe)

lm_fit = fit(lm_wflow, bike)

summary(lm_fit$fit$fit$fit)
```

because we set hr to be a factor, there are 24 hours, and therefore 23 dummy variables.

hr24 is the base model because it is not shown. it basically has a b_n of 0

hr 17 has the highest number of rides with a max slope of 408

```{r Question 6}
ggplot(bike, aes(x=temp, y=count)) + geom_point() + geom_smooth(method = lm) + theme_bw()
```

```{r Question 7}
bike_recipe2 = recipe(count ~ hr + temp, bike)
lm_model2 = 
  linear_reg() %>%
  set_engine("lm")

lm_wflow2 = 
  workflow() %>%
  add_model(lm_model2) %>%
  add_recipe(bike_recipe2)

lm_fit2 = fit(lm_wflow2, bike)

summary(lm_fit2$fit$fit$fit)
```


```{r Question 9}
bike_recipe3 = recipe(count ~ atemp + temp, bike)
lm_model3 = 
  linear_reg() %>%
  set_engine("lm")

lm_wflow3 = 
  workflow() %>%
  add_model(lm_model3) %>%
  add_recipe(bike_recipe3)

lm_fit3 = fit(lm_wflow3, bike)

summary(lm_fit3$fit$fit$fit)
```

```{r backward}
allmod = lm(count ~., bike)
backmod = stepAIC(allmod, direction = "backward", trace = TRUE) #trace = TRUE shows how the model is built (which variables are removed)
summary(backmod)
```


