# BAN 502 Predictive Analytics

## Module 4 Assignment 3 Random Forests

### Matthew Field


```{r load packages}
library("tidyverse")
library("tidymodels")
library("caret")
library("gridExtra")
library("vip")
library("ranger")
library(skimr)
drug <- read_csv("drug_data-2.csv")
```

Data cleaning

```{r cleaning}
#adding titles

names(drug) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity",
"Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive",
"SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis",
"Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh",
"LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")

#changing when use was to yes or no

drug[drug == "CL0"] = "No"
drug[drug == "CL1"] = "No"
drug[drug == "CL2"] = "Yes"
drug[drug == "CL3"] = "Yes"
drug[drug == "CL4"] = "Yes"
drug[drug == "CL5"] = "Yes"
drug[drug == "CL6"] = "Yes"

#factor recoding

drug_clean = drug %>% 
  mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>%
  mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44", "45_54",
                                      "55_64", "65_"))) %>%
  mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
  mutate(Education = factor(Education, labels = c("Under16", "At16", 
                                                  "At17", "At18",
                                                  "SomeCollege",
                                                  "ProfessionalCert", 
                                                  "Bachelors", "Masters", 
                                                  "Doctorate"))) %>%
  mutate(Country = factor(Country, labels = c("USA", "NewZealand", 
                                              "Other", "Australia",
                                              "Ireland","Canada","UK"))) %>%
  mutate(Ethnicity = factor(Ethnicity, labels = c("Black", "Asian", "White",
                                                  "White/Black", "Other",
                                                  "White/Asian", 
                                                  "Black/Asian"))) %>%
  mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>%
  select(-ID)

str(drug_clean)
```


we are focusing on nicotine, so I will remove all other drugs

```{r focus on nicotine}
drug_clean = drug_clean %>% select(!(Alcohol:Mushrooms)) %>% select(!(Semer:VSA))

```

```{r checking data set}
summary(drug_clean)
skim(drug_clean)

```

There is no missing data in the dataset

now can start going to work, start by splitting,

```{r}
set.seed(1234) 
drug_split = initial_split(drug_clean, prop = 0.7, strata = Nicotine) #70% in training
train = training(drug_split)
test = testing(drug_split)
```

1318 are in training set

visualization
```{r}
p1 = ggplot(train, aes(x = Age, fill = Nicotine)) + 
  geom_bar(position = "fill")
p2 = ggplot(train, aes(x = Impulsive, fill = Nicotine)) + 
  geom_bar(position = "fill")
p3 = ggplot(train, aes(x = Gender, fill = Nicotine)) + 
  geom_bar(position = "fill")
p4 = ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + 
  geom_bar(position = "fill")
grid.arrange(p1,p2,p3,p4)
```

You are less likely to smoke as you get older, so true

higher impulsiveness are also more likely to smoke, so true.


Set up our folds for cross-validation  
```{r}
set.seed(123)
rf_folds = vfold_cv(train, v = 5)
```


Random forest with an R-defined tuning grid (this model took about 5 minutes to run)
```{r}
drug_recipe = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>% #add tuning of mtry and min_n parameters
  #setting trees to 100 here should also speed things up a bit, but more trees might be better
  set_engine("ranger", importance = "permutation") %>% #added importance metric
  set_mode("classification")

drug_wflow = 
  workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(drug_recipe)

set.seed(123)
rf_res = tune_grid(
  drug_wflow,
  resamples = rf_folds,
  grid = 10 #try 10 different combinations of the random forest tuning parameters
)
```

Refining the parameters  
```{r}
drug_recipe2 = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model2 = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>% #add tuning of mtry and min_n parameters
  #setting trees to 100 here should also speed things up a bit, but more trees might be better
  set_engine("ranger", importance = "permutation") %>% #added importance metric
  set_mode("classification")

drug_wflow2 = 
  workflow() %>% 
  add_model(rf_model2) %>% 
  add_recipe(drug_recipe2)

rf_grid2 = grid_regular(
  mtry(range = c(2, 8)), #these values determined through significant trial and error
  min_n(range = c(5, 20)), #these values determined through significant trial and error
  levels = 10
)

set.seed(123)
rf_res_tuned = tune_grid(
  drug_wflow2,
  resamples = rf_folds,
  grid = rf_grid2 #use the tuning grid
)
```


An alternate view of the parameters  
```{r}
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  mutate(min_n = factor(min_n)) %>%
  ggplot(aes(mtry, mean, color = min_n)) +
  geom_line(alpha = 0.5, size = 1.5) +
  geom_point() +
  labs(y = "Accuracy")
```



select the best parameters
```{r}
best_rf = select_best(rf_res_tuned, metric="accuracy")

final_rf = finalize_workflow(
  churn_wflow,
  best_rf
)

final_rf
```
```{r}
#fit the finalized workflow to our training data
final_rf_fit = fit(final_rf, train)
```

Predictions  
```{r}
trainpredrf = predict(final_rf_fit, train)
head(trainpredrf)
```

Confusion matrix
```{r}
confusionMatrix(trainpredrf$.pred_class, train$Nicotine, 
                positive = "Yes")
```
Predictions on test
```{r}
testpredrf = predict(final_rf_fit, test)
head(testpredrf)
confusionMatrix(testpredrf$.pred_class, test$Nicotine, 
                positive = "Yes")
```

Check out variable importance
```{r}
final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")
```